\section{Conclusions}
\label{sec:conclusions}

We have presented \ROBERT{}, a pre-release Python framework for testable and
reproducible exoplanet atmospheric retrievals. Its principal contribution is a
set of explicit boundaries between observations, atmospheric state, opacity,
radiative transfer, instrument response, likelihoods, and inference. The
current implementation includes correlated-$k$ gas opacity and random overlap,
thermal emission and transmission, clear and scattering-cloud calculations,
one- and two-region daysides, multiple observation data sets, optimal
estimation, and UltraNest.

Controlled comparisons with PICASO already test the SH4 multiple-scattering
path at high numerical precision, while petitRADTRANS comparisons and the
WASP-69b data workflow broaden validation to molecular opacity, transmission,
and real multi-instrument observations. \roberttodo{Update this sentence after
all tagged benchmark artefacts and WASP-69b retrievals are complete.}

The WASP-80b analysis is designed as both a scientific reanalysis and a test of
the complete reproducibility chain. \roberttodo{Insert the main WASP-80b
findings and state clearly where they agree or disagree with published
constraints.}

Near-term development will focus on science-grade cloud microphysics,
high-stream scattering references, broader opacity validation, richer
disequilibrium chemistry, and robust long-run posterior workflows. Longer-term
interfaces permit reflected light, phase curves, multidimensional atmospheres,
line-by-line opacity, additional samplers, and multi-instrument covariance,
without requiring those capabilities to be claimed by the present release.

