\section{WASP-80b reanalysis}
\label{sec:results}

This section defines the analysis that will provide the principal science
result of the paper. It intentionally contains no inferred values before the
retrieval and robustness suite is complete.

\subsection{Data and model hierarchy}

We will reanalyse the published NIRCam/F322W2, NIRCam/F444W, and MIRI/LRS
$2.4$--$12\,\mu$m dayside emission spectrum of WASP-80b
\citep{Bell2023,WASP80Panchromatic2025}. The reduction, wavelength bins,
uncertainties, covariance assumptions, and any inter-instrument offsets will be
recorded as separate observation data sets rather than concatenated into an
anonymous spectrum. \roberttodo{Acquire the exact machine-readable spectrum,
record its DOI/MAST identifiers and checksums, and determine whether the
published covariance is available.}

The planned hierarchy is:
\begin{enumerate}
 \item a clear one-region free-chemistry model;
 \item a clear one-region equilibrium-chemistry model parameterized by
 metallicity and C/O;
 \item grey and microphysical cloud extensions;
 \item diluted and two-region dayside models; and
 \item targeted disequilibrium or quench-pressure extensions if justified by
 posterior predictive residuals.
\end{enumerate}
Each comparison will hold opacity, spectral response, priors on shared
parameters, and convergence criteria fixed where scientifically possible.

\subsection{Prespecified reported quantities}

The primary quantities are metallicity, C/O, thermal structure, CH$_4$ and
H$_2$O abundance constraints, cloud optical properties, and the evidence of
each model relative to the clear one-region baseline. We will report credible
intervals, prior-to-posterior contraction, posterior correlations, contribution
functions, and posterior predictive spectra. Molecular evidence will be tested
with nested model comparisons or abundance-posterior diagnostics declared
before inspecting the final result; a Gaussian ``sigma'' equivalent will not
be quoted without stating its definition.

\subsection{Robustness tests}

Robustness tests will include NIRCam-only and MIRI-only retrievals, alternative
temperature parameterizations, opacity-table resolution and bounds checks,
cloud-model sensitivity, removal of instrument offsets, repeated nested-sampling
runs, and injection--recovery at representative posterior points. The
published WASP-80b posterior products will be passed through the same plotting
and comparison code where their parameterization permits.

\subsection{Results completion gate}

Before this section is converted from a plan into results, all final runs must
have: (i) a versioned configuration and manifest; (ii) converged evidence and
posterior diagnostics; (iii) reproducible figures and tables; (iv) an archived
machine-readable posterior; and (v) an internal check that the text distinguishes
data-supported conclusions from model-dependent interpretation.

\roberttodo{Replace this paragraph with the quantitative results and add: data
figure; best-fitting/posterior-predictive spectra; abundance and thermal-profile
figure; corner or information-content summary; evidence table; and direct
comparison with \citet{WASP80Panchromatic2025}.}

