\documentclass[fleqn,usenatbib]{mnras}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}

\newcommand{\ROBERT}{\textsc{Robert}}
\newcommand{\roberttodo}[1]{\textcolor{magenta}{[TODO: #1]}}
\newcommand{\code}[1]{\texttt{#1}}

\title[ROBERT]{ROBERT: a reproducible retrieval framework for JWST exoplanet emission spectra}

\author[J. Taylor et al.]{
Jake Taylor$^{1}$\thanks{E-mail: \roberttodo{corresponding-author email}}
and \roberttodo{co-authors}\\
$^{1}$\roberttodo{affiliation}
}

\date{Accepted XXX. Received YYY; in original form ZZZ}
\pubyear{2026}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

\begin{abstract}
Atmospheric retrievals translate exoplanet spectra into constraints on thermal
structure, composition, clouds, and planetary properties, but scientific
interpretation can depend on choices made across opacity preparation,
radiative transfer, inference, and data handling. We present \ROBERT{}, an
open, typed Python framework developed for reproducible retrievals of JWST
exoplanet emission spectra. \ROBERT{} separates atmospheric state construction,
correlated-$k$ opacity preparation, radiative transfer, instrument response,
likelihood evaluation, and inference behind explicit interfaces, and records
the resulting configuration and data provenance. The current implementation
supports clear and cloudy thermal emission, transmission, multiple one-dimensional
regions, multiple data sets, optimal estimation, and nested sampling with
UltraNest. We validate individual numerical components and compare thermal
radiative transfer with PICASO and molecular emission and transmission with
petitRADTRANS. We further exercise the end-to-end data model using the
panchromatic JWST emission spectrum of WASP-69b. \roberttodo{Replace this final
sentence with the quantitative WASP-80b results, including retrieved metallicity,
C/O, molecular detections, and model evidences.} This work establishes a
testable route from retrieval-code development to a reproducible scientific
analysis while retaining clear boundaries between validated capability and
future physics.
\end{abstract}

\begin{keywords}
planets and satellites: atmospheres -- planets and satellites: gaseous planets
-- methods: numerical -- radiative transfer -- software: development
\end{keywords}

\input{sections/introduction}
\input{sections/methods}
\input{sections/results}
\input{sections/conclusions}
\input{sections/reproducibility}

\section*{Acknowledgements}
\roberttodo{Funding, computing facilities, and contributors.} The development
and writing workflow used OpenAI Codex. The authors reviewed and take
responsibility for all code, analyses, citations, and text produced with AI
assistance.

\section*{Data Availability}
The WASP-69b spectrum used for validation is distributed with source provenance
in the \ROBERT{} repository and originates from the machine-readable tables of
\citet{Schlawin2024}. \roberttodo{Add the WASP-80b data products, MAST programme
identifier, exact reduction version, retrieval posterior archive, and Zenodo
DOI.} Source code and analysis scripts will be archived at a versioned DOI on
acceptance; the development repository is \roberttodo{insert public GitHub URL}.

\bibliographystyle{mnras}
\bibliography{references}

\label{lastpage}
\end{document}

